---
title: "Lab 6"
author: "Jon Beeler"
date: "March 9, 2017"
output: github_document
html_document:
  df_print: paged
  keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = F, warning = F)
library(RCurl)
library(ggmap)
library(jsonlite)
library(memisc)
library(censusapi)
library(sp)
library(maps)
library(maptools)
library(dplyr)
library(raster)
library(rgeos)
library( rgdal )
library(V8)
library( geojsonio )

```
Part 1: 
1. Unclipped: Onondaga County census tracts and primary and secondary roads. Plotted together, highlighting interstates
```{r}
setwd( "/Users/beelerenator/Documents/Graduate School/MPA Syracuse/DDMII/Lab 6")

#read in shape file of onondaga county
onondoga <- readShapePoly( fn="tl_2010_36067_tract10", proj4string=CRS("+proj=longlat +datum=WGS84") )
roads <- readShapeLines( fn="tl_2015_36_prisecroads", proj4string=CRS("+proj=longlat +datum=WGS84") )

dat.onondoga <- as.data.frame( onondoga )

par( mar=c(0,0,0,0) )
plot( onondoga,  border="gray10" )

### ADD ROADS
#download.file("ftp://ftp2.census.gov/geo/tiger/TIGER2015/PRISECROADS/tl_2015_36_prisecroads.zip", "primary_and_secondary_roads.zip" )
#unzip( "primary_and_secondary_roads.zip" )
#file.remove( "primary_and_secondary_roads.zip" )

dat.roads <- as.data.frame( roads )
table( roads$RTTYP )

### Extract Interstates

par( mar=c(0,0,0,0) )
plot( onondoga,  border="gray80" )
plot( roads, col="steelblue", lwd=2, add=T )

interstate <- roads[ roads$RTTYP == "I" , ]

par( mar=c(0,0,0,0) )
#y.max <- (43.26848)
#y.min <- (42.77941)
#x.max <- (-75.8487)
#x.min <- (-76.46031)

plot( onondoga,  border="gray80")
plot( roads, col="steelblue", lwd=2, add=T )
plot( interstate, col="red", add=T )

```
2. and 3.  Clipped census tracts with interstate extacted
```{r}

b <- bbox(onondoga)

gClip <- function(shp, bb) {
if(class(bb) == "matrix") b_poly <- as(extent(as.vector(t(bb))), "SpatialPolygons") 
else b_poly <- as(extent(bb), "SpatialPolygons")
gIntersection(shp, b_poly, byid = T)
}

onondoga_clipped <- gClip(onondoga, b)
roads_clipped <- gClip(roads, b)
interstate_clipped <- gClip(interstate, b)

plot(onondoga_clipped, border="gray80")
plot(roads_clipped, col="steelblue", lwd=2, add=T )
plot( interstate_clipped, col="red", add=T )

```

4. Buffer from interstate 
```{r}

setwd( "/Users/beelerenator/Documents/Graduate School/MPA Syracuse/DDMII/Lab 6")
housing.census.crime <- read.csv("housing.census.crime.dat.csv", stringsAsFactors = F)

plot(onondoga_clipped, border="gray80")
plot( interstate_clipped, col="red", add=T )
points(housing.census.crime$lon, housing.census.crime$lat, pch=19, cex=.2)
buff1 <- gBuffer(interstate_clipped, width=.005, byid=TRUE )
plot(buff1, add=T )

```
5. New categorical variable indiacting whether houses fall in interstate buffer
```{r}
buff.dissolved <- gUnaryUnion( buff1, id = NULL )
lat.long <- dplyr::select(housing.census.crime, lon, lat)

lat.long <- SpatialPoints(lat.long, proj4string=CRS("+proj=longlat +datum=WGS84") )
over.buffer <- over(lat.long, buff.dissolved)

over.buffer <- as.data.frame(over.buffer)
over.buffer[is.na(over.buffer)] <- 0
over.buffer

housing.census.crime.hway.buffer.interstate <- cbind(housing.census.crime, over.buffer)

housing.census.crime.hway.buffer.interstate <- plyr::rename(housing.census.crime.hway.buffer.interstate, c("over.buffer"="Near Interstate"))

head(housing.census.crime.hway.buffer.interstate)
```

Part 2: Use Syracuse parcel data to create a buffer from industrial zones 
```{r}

setwd("/Users/beelerenator/Documents/Graduate School/MPA Syracuse/DDMII/Lab 6")
syr.1 <- readOGR("syr_parcels.geojson")

#syr <- readShapePoly( fn="01-05-2015", proj4string=CRS("+proj=longlat +datum=WGS84") )
syr1.dat <- as.data.frame(syr.1)

unique(syr1.dat$LandUse)

industrial <- syr.1[syr.1$LandUse == "Industrial", ]
not.industrial <- syr.1[syr.1$LandUse != "Industrial", ]

plot( not.industrial, col="gray80", border=F)
plot( industrial, col="red", border=F, add=T )

buff.industrial <- gBuffer(industrial, width = .003621, capStyle = "FLAT", quadsegs = 1)

plot(buff.industrial, add=T)

```
2. Map Houses within Industrial Buffer and create variable indicating whether houses fall within industrial buffer
```{r}

  plot( not.industrial, col="gray80", border=F)
  plot( industrial, col="red", border=F, add=T )
  points(housing.census.crime$lon, housing.census.crime$lat, pch=19, cex=.2, col="darkorchid4")
  plot(buff.industrial, add=T)
  buff.industrial.dissolved <- gUnaryUnion( buff.industrial, id = NULL )


lat.long <- dplyr::select(housing.census.crime, lon, lat)
lat.long <- SpatialPoints(lat.long, proj4string = CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0") )

over.buffer.landuse <- over(lat.long, buff.industrial.dissolved)

over.buffer.landuse <- as.data.frame(over.buffer.landuse)
over.buffer.landuse[is.na(over.buffer.landuse)] <- 0
over.buffer.landuse

housing.census.crime.hway.interstate.industrial.buffers <- cbind(housing.census.crime.hway.buffer.interstate, over.buffer.landuse)

housing.census.crime.hway.interstate.industrial.buffers <- plyr::rename(housing.census.crime.hway.interstate.industrial.buffers, c("over.buffer.landuse"="Near Industrial Site"))

head(housing.census.crime.hway.interstate.industrial.buffers)

```
3. Create a buffer from schools
```{r}

schools <- syr.1[syr.1$LandUse == "Schools", ]
not.schools <- syr.1[syr.1$LandUse != "Schools", ]

plot( not.schools, col="gray80", border=F)
plot( schools, col="lightseagreen", border=F, add=T )

buff.schools <- gBuffer(schools, width = .0018105, capStyle = "FLAT", quadsegs = 1)
plot(buff.schools, add=T)

```
4. Map Houses within School Buffer and create variable indicating whether houses fall within school buffer
```{r}
  plot( not.schools, col="gray80", border=F)
  plot( schools, col="lightseagreen", border=F, add=T )
  points(housing.census.crime$lon, housing.census.crime$lat, pch=19, cex=.2, col="darkorchid4")
  plot(buff.schools, add=T)
  buff.schools.dissolved <- gUnaryUnion( buff.schools, id = NULL )


lat.long <- dplyr::select(housing.census.crime, lon, lat)
lat.long <- SpatialPoints(lat.long, proj4string = CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0") )

over.buffer.schools <- over(lat.long, buff.schools.dissolved)

over.buffer.schools <- as.data.frame(over.buffer.schools)
over.buffer.schools[is.na(over.buffer.schools)] <- 0
over.buffer.schools

housing.census.crime.hway.interstate.industrial.school.buffers <- cbind(housing.census.crime.hway.interstate.industrial.buffers, over.buffer.schools)

housing.census.crime.hway.interstate.industrial.school.buffers <- plyr::rename(housing.census.crime.hway.interstate.industrial.school.buffers, c("over.buffer.schools"="Near Schools"))

head(housing.census.crime.hway.interstate.industrial.school.buffers)

```

